---
title: 'Homework #11'
author: "Brendan Case"
date: "10/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Advanced ggplotting

```{r, results=F}
library(ggplot2)
library(ggmosaic)
library(GGally)
library(ggthemes)
```


Two useful extentions of `ggplot2` that I found this semester were the `ggmosaic` and `GGally` packages. I'll include an example of using each here, using the `iris` dataset. As for themes, I personally really like the `theme_few` theme, especially for including in mathematical texts. The clear background and slight inlay looks really good, and I like the generally softer colors.

```{r, message=F}
# make another (silly) categorical variable:
dat = cbind(iris, comp=iris$Sepal.Length < iris$Petal.Length + iris$Petal.Width)
head(dat)
p = ggplot(dat) + theme_few() + scale_color_few()
p + geom_mosaic(aes(x = product(Species), fill=Species))
p + geom_mosaic(aes(x = product(Species, comp), fill=Species))
```

`GGally` is especially useful for viewing relationships between variables and for visualizing model fitting. I find the `ggpairs` function to make much more pretty results than base R's `pairs` function, though it is easy to have information overload.

```{r, fig.width=9, fig.height=7, message=F}
ggpairs(iris, mapping=aes(color=Species)) + theme_few()
```


