---
title: 'Homework #10'
author: "Brendan Case"
date: "10/24/2018"
output: html_document
---

# Randomization Tests

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

For simplicity, I'm just going to use the fake data which was produced in Homework 7/8 for this project. Namely, I generate several response variables from a normal distribution with different means and variances. Since this data has three response variables, I will not be doing a regression test like we did in class, but instead a randomization version of a difference in means test.

```{r}
##################################################
# FUNCTION: eab_sim
# generate data simulating the number of survining ash trees for different
# treatment cases
# input: parameters for the data
# output: data frame (long format)
#------------------------------------------------- 
eab_sim = function(n=1000,
                   groups = c('control', 'quar', 'aware'),
                   means = c(400, 650, 600),
                   std_devs = c(200, 50, 150)) {
  dat = data.frame(rnorm(n, means[1], std_devs[1]),
                   rnorm(n, means[2], std_devs[2]),
                   rnorm(n, means[3], std_devs[3]))
  colnames(dat) = groups
  return(dat)
}
```


```{r}
####### ###########################################
# function: readData
# read in (or generate) data set for analysis
# input: none
# output: 3 column data frame of observed data
#------------------------------------------------- 
readData <- function() {
  return(eab_sim())
}

##################################################
# function: shuffleData
# randomize data by randomly swapping 1/3 of the data between groups
# input: 3-column data frame
# output: 3-column data frame
#------------------------------------------------- 
shuffleData <- function(data) {
  temp = data
  s1 = sample(1:(length(data[,1])/3), replace=F)
  s2 = sample(1:(length(data[,1])/3), replace=F)
  s3 = sample(1:(length(data[,1])/3), replace=F)
  data[,1] = c(data[-s1,1], temp[s1,2])
  data[,2] = c(data[-s2,2], temp[s2,3])
  data[,3] = c(data[-s3,3], temp[s3,1])
  return(data)
}

##################################################
# function: getPVal
# calculate p value from simulation
# input: list of observed metric, and vector of simulated metrics
# output: lower, upper tail probability values
#------------------------------------------------- 
getPVal <- function(z=NULL) {
                    if(is.null(z)){
                      z <- list(xObs=runif(1),xSim=runif(1000))}
pLower <- mean(z[[2]]<=z[[1]])
pUpper <- mean(z[[2]]>=z[[1]])
return(c(pL=pLower,pU=pUpper))
}

##################################################
# function: plotRanTest
# create ggplot of histogram of simulated values
# input: list of observed metric and vector of simulated metrics
# output: saved ggplot graph
#------------------------------------------------- 
plotRanTest <- function(z=NULL) {
                if(is.null(z)){
                  z <- list(rnorm(1),rnorm(1000)) }
dF <- data.frame(ID=seq_along(z[[2]]),simX=z[[2]])
p1 <- ggplot(data=dF,mapping=aes(x=simX))
p1 + geom_histogram(mapping=aes(fill=I("goldenrod"),color=I("black"))) +
geom_vline(aes(xintercept=z[[1]],col="blue")) 

}

set.seed(1234)
df = readData()
df_rand = shuffleData(df)
# what is the probability that the observed data came from the shuffled data?
```

## 2

Normally, one would do two-sample t-tests for this type of thing, but I already wrote a function to do that in Homework 8.
