---
title: 'Homework #10'
author: "Brendan Case"
date: "10/24/2018"
output: html_document
---

# Randomization Tests

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## 1

For simplicity, I'm just going to use the fake data which was produced in Homework 7/8 for this project. Namely, I generate several response variables from a normal distribution with different means and variances. Since this data has three response variables, I will not be doing a regression test like we did in class, but instead a randomization version of a difference in means test based on a control and response group. My control group will be the number of healthy ash trees after no intervention, and my response group will be the number of healthy ash trees after a quarantine. The p-value then becomes the probability that the observed difference in means between the control and response group came from the distribution of mean differences in the shuffled data.

```{r}
##################################################
# FUNCTION: eab_sim
# generate data simulating the number of survining ash trees for different
# treatment cases
# input: parameters for the data
# output: data frame (long format)
#------------------------------------------------- 
eab_sim = function(n=1000,
                   groups = c('control', 'quar'),
                   means = c(400, 430),
                   std_devs = c(200, 150)) {
  dat = data.frame(rnorm(n, means[1], std_devs[1]),
                   rnorm(n, means[2], std_devs[2]))
  colnames(dat) = groups
  return(dat)
}
```


```{r}
####### ###########################################
# function: readData
# read in (or generate) data set for analysis
# input: none
# output: 3 column data frame of observed data
#------------------------------------------------- 
readData <- function() {
  return(eab_sim(groups=c('control', 'quar'), means=c(400, 410),
                 std_devs=c(200, 150)))
}

##################################################
# function: shuffleData
# randomize data by randomly swapping 1/2 of the data between groups
# input: 2-column data frame
# output: shuffled 2-column data frame
#------------------------------------------------- 
shuffleData <- function(data) {
  data = df
  temp = data
  s = sample(1:(length(data[,1])), length(data[,1])/2, replace=F)
  data[,1] = c(data[-s,1], temp[s,2])
  data[,2] = c(data[-s, 2], temp[s, 1])
  # plot(data[,1])
  return(data)
}

##################################################
# function: getPVal
# calculate p value from simulation
# input: list of observed metric, and vector of simulated metrics
# output: lower, upper tail probability values
#------------------------------------------------- 
getPVal <- function(null_dist, obs_mean, n=1000){
  pLower <- length(null_dist[null_dist > obs_mean]) / n
  return(pLower)
}

##################################################
# function: getMetric
# calculate metric for randomization test
# input: 2-column data frame
# output: null distribution
#------------------------------------------------- 
getMetric <- function(data, n=1000) {
  ret = rep(-1, 1000)
  for (i in 1:n) {
    shuf = shuffleData(data)
    ret[i] = mean(shuf[,1]) - mean(shuf[,2])
  }
  return(ret)
}

##################################################
# function: plotRanTest
# create ggplot of histogram of simulated values
# input: list of observed metric and vector of simulated metrics
# output: saved ggplot graph
#------------------------------------------------- 
plotRanTest <- function(null_dist, obs_mean) {
  ggplot(data.frame(null_dist), aes(null_dist)) + geom_histogram(bins=30) +
    geom_vline(xintercept=obs_mean, col='red', linetype='dashed')
}

## program body
set.seed(1234)
df = readData()
null_dist = getMetric(df)
getPVal(null_dist, mean(df$quar) - mean(df$control))
plotRanTest(null_dist, mean(df$quar) - mean(df$control))
```

## 2

Normally, one would do two-sample t-tests for this type of thing, but I already wrote a function to do that in Homework 8.
