---
title: 'Homework #9'
author: "Brendan Case"
date: "10/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Looping Structures

Here are the first 3 exercises from Homework 10. I put them here so that all the looping exercises are contained in the same place.

## 1

```{r}
numzeros = function(x = c(1)) {
  if (!is.numeric(x)) {
    print("WARNING: non-numeric vector given")
    return(0)
  }
  counter = 0
  for (x_i in x) {
    if (x_i == 0 ) {
      counter = counter + 1
    }
  }
  return(counter)
}
```

## 2

Mine is longer than a single line to deal with invalid arguments.

```{r}
numzeros2 = function(x = c(1)) {
  if (!is.numeric(x)) {
    print("WARNING: non-numeric vector given")
    return(0)
  }
  return(length(x[x == 0]))
}
```

## 3

```{r}
q3 = function(nrow=0, ncol=0) {
  ret = matrix(NA, nrow, ncol)
  for (i in 1:nrow) {
    for (j in 1:ncol) {
      ret[i, j] = i * j
    }
  }
  return(ret)
}
```

## Old exercises:

These were the excercises from last week, that were then removed. For completeness, I figured I would include them.

### 3

```{r}
maxdiff = function(x = c(0)) {
  if (!is.numeric(x)) {
    print("WARNING: non-numeric vector given")
    return(0)
  }
  diffs = c()
  for (i in x) {
    for (j in x) {
      diffs = c(diffs, abs(i - j))
    }
  }
  return(max(diffs))
}
```

### 4

Modify the above function:

```{r}
maxdiff = function(x = c(0)) {
  if (!is.numeric(x)) {
    print("WARNING: non-numeric vector given")
    return(0)
  }
  xis = integer(0)
  xjs = integer(0)
  diffs = integer(0)
  is = integer(0)
  js = integer(0)
  for (i in 1:length(x)) {
    for (j in 1:length(x)) {
        xis = c(xis, x[i])
        xjs = c(xjs, x[j])
        diffs = c(diffs, abs(x[i] - x[j]))
        is = c(is, i)
        js = c(js, j)
    }
  }
  df = data.frame(xis=xis, xjs=xjs, diffs=diffs, is=is, js=js)
  first = df[df["diffs"] == max(df["diffs"]),][1,]
  return(list(c(first$xis, first$xjs),
              first$diff,
              c(first$is, first$js)))
}
maxdiff(c(-3, 0 ,1, 3))
```

### 5

```{r}
maxdiff2 = function(x = c(0)) {
  if (!is.numeric(x)) {
    print("WARNING: non-numeric vector given")
    return(0)
  }
  temp = 0
  for (i in x) {
    for (j in x) {
      if (temp < abs(i - j)) {
        temp = abs(i - j)
      }
    }
  }
  return(temp)
}
```

### 6

```{r}
matmut = function(A=matrix(), B=matrix()) {
  if (ncol(A) != nrow(B)) {
    print("WARNING: incorrect matrix dimensions")
    return(NULL)
  }
  n = nrow(A)
  p = ncol(B)
  ret = matrix(NA, nrow=n, ncol=p)
  for (i in 1:n) {
    for (j in 1:p) {
      ret[i,j] = sum(A[i,] * B[,j])
    }
  }
  return(ret)
}